---
title: "Assessing whether intra and inter-genomic variation hinder the utility of ASVs"
author: "Vaibhav Gawde"
date: "2023-06-20"
output: 
  github_document:
    html_preview: false
---


## Questions
* Within a genome, how many distinct sequences of the 16s rRNA gene are present relative to the number of copies per genome? How far apart are these sequences from each other? How does this scale from genome to kingdoms?
* Within a taxa (any level), how may ASVs from that taxa are shared with sister taxa? How does this change with taxonomic level? Variable region?

* Make sure we have taxonomic data for all our genomes
* Read FASTA files into R 
* inner_join with tsv file
* group_by / summarize to count number of sequences and copies per genome

```{r library_calls, message=FALSE, echo=FALSE}
library(tidyverse)
library(data.table)
library(rmarkdown)
```

### Dependencies:
* [mothur v1.46.1](https://github.com/mothur/mothur/releases/tag/v1.46.1) - `code/install_mothur.sh` installs mothur
* `wget`
* `r R.version.string` 
  * `tidyverse` (v. `r packageVersion("tidyverse")`)
  * `data.table` (v. `r packageVersion("data.table")`)
  * `rmarkdown` (v. `r packageVersion("rmarkdown")`)
  
### My computer

```{r sessioninfo, echo=FALSE}
sessionInfo()
```