---
title: "Analyzing the sensitivity and specificity of ASVs for discriminating between genomes"
author: "Vaibhav Gawde"
date: "2023-06-21"
output: 
  github_document:
  html_preview: true
---

```{r, message=FALSE}
library(tidyverse)
library(here)
```

### Need to determine the number of *rrn* operons accross the v34 regions of the genomes

```{r}
count_tibble <- read_tsv(here("data/processed/rrnDB.count_tibble"), 
                         col_types = "cccd")
```

We want to count and plot the number of copies per genome
We'll maybe do this for the **v4** sub-region as an example

```{r}
count_tibble %>% 
  filter(region == "v4") %>% 
  group_by(genome) %>% 
  summarise(n_rrn = sum(count), .groups = "drop") %>% 
  ggplot(aes(x=n_rrn)) + geom_histogram(binwidth = 1)

count_tibble %>% 
  filter(region == "v4") %>% 
  group_by(genome) %>% 
  summarise(n_rrn = sum(count)) %>% 
  count(n_rrn) %>% 
  mutate(fraction = n / sum(n))
```
We see that most genomes actually have more than one copy of the *rrn* operon. Are those copies the same sequence / ASV..?

### Determine number of ASVs per genome

Considering most genomes have multiple copies of the *rrn* operon, we need to know whether they are the same ASV. Otherwise we run the risk of splitting a single genome into multiple ASVs.

Now we'll take into account all the sub-regions that we've extracted ie v4, v34, v45
```{r}
count_tibble %>% 
  group_by(region, genome) %>% 
  summarise(n_asv = n(), n_rrn = sum(count)) %>% 
  group_by(region, n_rrn) %>% 
  summarise(med_n_asv = median(n_asv),
            mean_n_asv = mean(n_asv),
            lq_n_asv = quantile(n_asv, prob = 0.25),
            uq_n_asv = quantile(n_asv, prob = 0.75)) %>% 
  filter(n_rrn == 7)


count_tibble %>% 
  group_by(region, genome) %>% 
  summarise(n_asv = n(), n_rrn = sum(count)) %>% 
  ggplot(aes(x=n_rrn, y=n_asv, color=region)) + geom_smooth(method="lm")
```

We can see that the sub-region **v34** has more number of ASVs per genome than **v4** and **v45**. **v4** is common between them so it just says that the **v3** region has more number of unique ASVs than **v5** region.

### Determine whether an ASV is unique to genomes they're found in

Instead of looking at the number of ASVs per genome, we want to see the number of genomes per ASV.

```{r}
count_tibble %>% 
  group_by(region, asv) %>% 
  summarise(n_genomes = n()) %>% 
  count(n_genomes) %>% 
  mutate(fraction = n/sum(n)) %>% 
  filter(n_genomes == 1)
```

We can see that with the these sub-region, how many % of the ASVs were unique to a genome