---
title: "Analyzing the sensitivity and specificity of ASVs for discriminating between genomes"
author: "Vaibhav Gawde"
date: "2023-06-21"
output: 
  github_document:
    html_preview: false
---

```{r, message=FALSE}
library(tidyverse)
library(here)
```

### Need to determine the number of *rrn* operons accross the v34 regions of the genomes

Our analysis will use only the v34 region of the sequences 

```{r}
v34 <- read_tsv(here("data/v34/rrnDB.count_tibble"))
```

We want to count and plot the number of copies per genome

```{r}
v34 %>% 
  group_by(genome) %>% 
  summarise(n_rrn = sum(count), .groups = "drop") %>% 
  ggplot(aes(x=n_rrn)) + geom_histogram(binwidth = 1)

v34 %>% 
  group_by(genome) %>% 
  summarise(n_rrn = sum(count), .groups = "drop") %>% 
  count(n_rrn) %>% 
  mutate(fraction = n / sum(n))
```
We see that most genomes actually have more than one copy of the *rrn* operon. Are those copies the same sequence / ASV..?

### Determine number of ASVs per genome

Considering most genomes have multiple copies of the *rrn* operon, we need to know whether they are the same ASV. Otherwise we run the risk of splitting a single genome into multiple ASVs.

```{r}
v34 %>% 
  group_by(genome) %>% 
  summarise(n_asv = n(), n_rrn = sum(count)) %>% 
  group_by(n_rrn) %>% 
  summarise(mean_n_asv = mean(n_asv),
            lq_n_asv = quantile(n_asv, prob = 0.25),
            uq_n_asv = quantile(n_asv, prob = 0.75)) 


v34 %>% 
  group_by(genome) %>% 
  summarise(n_asv = n(), n_rrn = sum(count)) %>% 
  ggplot(aes(x=n_rrn, y=n_asv)) + geom_smooth(method="lm")
```

We can see that the number of ASVs increase at a rate of about 5 ASV per 20 copies of *rrn* operon of the v34 region in the genome

### Determine whether an ASV is unique to genomes they're found in

Instead of looking at the number of ASVs per genome, we want to see the number of genomes per ASV.

```{r}
v34 %>% 
  group_by(asv) %>% 
  summarise(n_genomes = n()) %>% 
  count(n_genomes) %>% 
  mutate(fraction = n/sum(n))
```
We see that with the v34 region, almost 78% of the ASVs were unique to a genome

### Now we'll see that if the sensitivity and specificity changes if we look at even shorter region of the genome, only the v4 region

We know that the v4 region is less diverse than the v34. So does the number of ASVs per genome differ than for v34? 

```{r}
v4 <- read_tsv(here("data/v4/rrnDB.count_tibble"))

v4 %>% 
  group_by(genome) %>% 
  summarise(n_rrn = sum(count), .groups = "drop") %>% 
  ggplot(aes(x=n_rrn)) + geom_histogram(binwidth = 1)

v4 %>% 
  group_by(genome) %>% 
  summarise(n_rrn = sum(count), .groups = "drop") %>% 
  count(n_rrn) %>% 
  mutate(fraction = n / sum(n))
```

Determining the number of ASVs per genome 
```{r}
v34 %>% 
  group_by(genome) %>% 
  summarise(n_asv = n(), n_rrn = sum(count)) %>% 
  group_by(n_rrn) %>% 
  summarise(mean_n_asv = mean(n_asv),
            lq_n_asv = quantile(n_asv, prob = 0.25),
            uq_n_asv = quantile(n_asv, prob = 0.75)) 


v34 %>% 
  group_by(genome) %>% 
  summarise(n_asv = n(), n_rrn = sum(count)) %>% 
  ggplot(aes(x=n_rrn, y=n_asv)) + geom_smooth(method="lm")
```


We can see that the number of ASVs increase at a rate of about 2.5 ASV per 20 copies of *rrn* operon of the v4 region in the genome


Now let's look at the specificity of an ASV for a genome


```{r}
v4 %>% 
  group_by(asv) %>% 
  summarise(n_genomes = n()) %>% 
  count(n_genomes) %>% 
  mutate(fraction = n/sum(n))
```
We find that with the v4 region, almost 76% of the ASVs were unique to a genome

### To be continued...

* Can we correct for over representation?
* Consider analysis at species, genus, family, etc. levels
* Consider a broader definition of ASVs
